Qualitative Probabilistic Programming

### 2015-02-25 paper/paper.tex
Update intro

  Probabilistic programming languages have made it easier to specify Bayesian
  models as programs.  Once a model is written as a probabilistic
  program, it is possible to use generic inference algorithms (such as
  Metropolis Hastings) to perform inference.

  In this paper, we discuss a feature of probabilistic programming languages
  that makes it easier to specify some models.
  Specifically, we allow the programmer to specify that some random functions
  are unknown.  The system will fill in these unknown functions with
  some reasonable default family of distributions based on the types and
  proceed to infer the parameters to this distribution using the
  expectation maximization algorithm.

  This feature has multiple advantages.  First, it is easier to write
  models without knowing about the class of models being used.  This should
  make probabliistic programming more accessible to non-experts.  Secondly,
  parameter inference is more efficient if specialized algorithms are used
  rather than the generic algorithms used to infer other random variables
  (such as Metropolis Hastings).

  We define an example probabilistic programming language with this feature
  (Quipp) and show how it can be used to write machine learning models
  concisely.

@@

  Inference in these models is performed using the expectation-maximization
  algorithm, with alternating steps of inferring latent variables and optimizing
  parameters.

  - Motivation for Quipp
    - Explanation of "unknown functions"
    - Writing machine learning algorithms as probabilistic programs
    - Accessibility to non-experts
    - Comparison to existing probabilistic programming languages
      - In other languages, use random variables for parameters
      - Random variables slower because they are updated independently

@@

    Let $\psi_x$ be a matrix such that $\psi \phi(x)$ is a vector of the
    features of $x$.  Then

>>>>

  By separating model specification and inference, probabilistic programming 
  has made it easier for non-experts to implement and use probabilistic models.
  Practitioners frequently have strong intuitions about the {\em structure}
  of their domain knowledge, such as which latent variables exist and what
  their causal relations are, and probabilistic programming allows them to encode
  this knowledge. However, it also requires them to specify the specific parametric
  shape  and parameterization of any distributions used, and intuitions tend to
  be much less precise there.
  We present Quipp, a system that does {\em not} require such specification;
  instead, random variables and random functions can be left undefined
  and will automatically be filled in under maximum entropy assumptions
  based on their types and available datasets.

  Our formalism can concisely express a wide variety of models that machine
  learning practitioners care about, and we provide an expectation
  maximization algorithm that can learn the parameters for many of
  these models with reasonable efficiency. This system makes it easy
  for non-experts to encode their beliefs about the data and to get
  predictions based on as few additional assumptions as possible.

  % This feature has multiple advantages.  First, it is easier to write
  % models without knowing about the class of models being used.  This should
  % make probabliistic programming more accessible to non-experts.  Secondly,
  % parameter inference is more efficient if specialized algorithms are used
  % rather than the generic algorithms used to infer other random variables
  % (such as Metropolis Hastings).

  % We define an example probabilistic programming language with this feature
  % (Quipp) and show how it can be used to write machine learning models
  % concisely.

@@

  % Inference in these models is performed using the expectation-maximization
  % algorithm, with alternating steps of inferring latent variables and optimizing
  % parameters.

  % - Motivation for Quipp
  %   - Explanation of "unknown functions"
  %   - Writing machine learning algorithms as probabilistic programs
  %   - Accessibility to non-experts
  %   - Comparison to existing probabilistic programming languages
  %     - In other languages, use random variables for parameters
  %     - Random variables slower because they are updated independently

  In the following, we first specify the syntax used to write Quipp programs,
  including the notation for unknown variables and functions.
  We describe the class of exponential family variables and functions that our system can learn,
  and present the expectation maximization algorithm used to learn them.
  We then demonstrate the expressiveness of our language, and the broad
  applicability of our algorithm, by writing some of the most common machine learning models
  in Quipp: clustering, factor analysis, a Hidden Markov model, Latent Dirichlet Allocation, and
  a neural net.

@@

    Let $\psi_x$ be a matrix such that $\psi \phi(x)$ is a vector x of the
    features of $x$.  Then
